---
title: "Geospatial Data Science - Final project"
author:
- Moritz Peist
- Noemi Lucchi
- Simon Vellin
subtitle: Analysis of US Road Accidents and Construction Sites (TBC)
output:
  html_document:
    toc: true
    df_print: paged
  html_notebook:
    df_print: paged
    toc: true
---

<hr>

```{css, echo=FALSE}
h1.title {
  text-align: center;
}

h4.author, h4.date {
  text-align: center;
}

.author {
  text-align: center;
}

.date {
  text-align: center;
}

/* Add styling for subtitle */
h3.subtitle {
  text-align: center;
  font-style: italic;
  margin-top: -10px;
  color: #606060;
}
```

```{r setup, include=FALSE}
# Disable warning messages
knitr::opts_chunk$set(warning = FALSE, message = FALSE)
```

<div style="text-align: center;">
  <img src="../Barcelona_School_of_Economics_logo.svg" alt="BSE logo">
</div>

<hr>

# Purpose of Study

This study investigates the causal effect of road construction and closures on accident rates in the United States by employing a propensity score matching framework to compute a difference-in-differences regression. Using two distinct datasets — US Accidents (2016–2023) and US Road Construction and Closures (2016–2021) — this paper aims to rigorously assess how construction-related disruptions influence the frequency and severity of traffic accidents.

Although a substantial body of literature has examined the latter, the specific integration of causal inference techniques remains less prevalent. As such, our approach provides a robust methodology by matching treated and control segments based on relevant covariates, which isolates the studied effect by comparing changes in accident rates over time between matched groups.

After quickly reviewing existing literature, we will justify the use of propensity score matching and difference-in-differences before implementing our code in Part 3. Finally, the last section will display our results while mentioning the limits of our study, as well as the potential improvements for subsequent analysis.

# Part 1 - Literature Review

In order to familiarize ourselves with the topic, we have been through two existing papers relating to accidents in the US, serving as a baseline understanding to start our study. Although we would have preferred introducing one of Mohamed Abdel-Aty's studies, widely recognized for his contributions to transportation safety, he employs advanced statistical modeling techniques which appears less related to Geospatial & Economic litterature compared to the ones that we decided to present below.

## 2.1 "Effects of truck traffic on crash injury severity on rural highways in Wyoming using Bayesian binary logit models" by Ahmed et al. (2018)

This paper explores the impact of truck traffic and roadway geometry on crash frequency within highway work zones, using a case study from Wyoming spanning 2012–2016.

Using crash data and statistical modeling, the authors assess how construction-related factors, like lane closures and heavy vehicle presence, elevate accident risks. Their paper reveals that increased truck volumes and sharper geometric features significantly amplify crash rates, with work zones acting as hotspots for such incidents. While the paper employs logistic regression rather than causal inference techniques, it underscores the real-world relevance of construction disruptions on traffic safety.

This work provides a practical foundation for our study, highlighting the need to isolate causal effects of construction using advanced methods. By focusing on observable crash patterns, Ahmed et al. set the stage for deeper methodological exploration in our difference-in-differences framework.

URL - https://www.sciencedirect.com/science/article/pii/S0001457518301507

## 2.2 "Comparison of characteristics between fatal and injury accidents in the highway construction zones" by Li and Bai (2008)

This paper investigates the characteristics of accidents in highway construction zones from a statistical perspective, comparing fatal and injury crashes using data from Utah between 1992 and 2004.

Using traffic volume, construction duration, and time of day, their results show that construction zones markedly increase accident likelihood, with fatal crashes tied to higher speeds and nighttime conditions. Although the study relies on descriptive statistics and regression rather than causal frameworks, it offers critical insights into how construction alters road safety dynamics.

It supports our research by establishing a clear link between construction activities and accident rates, justifying our use of propensity score matching and difference-in-differences. Ideally, our approach aims to address the methodological gaps they leave unexplored.

URL - https://www.sciencedirect.com/science/article/pii/S0925753507001014

# Part 2 - Methodology

In this section, we start by providing summary statistics of the datasets we have found for our study, while detailing the theoretical justifications underlying the two main tools used to compute our results.

## 2.1 Dataset Overview

For both datasets, the data was collected using multiple APIs such as the US and state departments of transportation, law enforcement agencies, traffic cameras, and traffic sensors within the road networks.

### a. US Accidents (2016 - 2023)

This is a countrywide car accident dataset that covers 49 states of the USA. The dataset currently contains approximately 7.7 million accident records.

```{Descriptive Statistics of Data 1}

```

### b. US Road Construction and Closures (2016 - 2021)

This is a countrywide dataset of road construction and closure events, which covers 49 states of the US, ranging from fixing pavements to substantial projects that could take months to finish. Currently, there are about 6.2 million construction and closure records in this dataset.

```{Descriptive Statistics of Data 2}

```

## 2.2 Propensity Score Matching

Propensity Score Matching (PSM) is a statistical technique designed to estimate causal effects by balancing treatment and control groups based on observed covariates, tackling selection bias inherent to observational data like ours.

Here, “treatment” denotes road segments with construction or closures, while “control” covers unaffected segments. PSM hinges on the Conditional Independence Assumption, which stipulates that given a set of covariates $X$ (e.g., traffic volume, road type, weather), treatment assignment is independent of the outcome—accident rates. This is expressed as:

$$Y(0), Y(1) \perp T \mid X,$$

where $Y(1)$ and $Y(0)$ represent potential outcomes under treatment and control, and $T$ is the treatment indicator. PSM estimates the propensity score, the probability of treatment given covariates, as:

$$PSM(X) = P(T = 1 \mid X),$$

typically via logistic regression. We then match treated and control units with similar $PSM(X)$ values, ensuring comparability. The average treatment effect on the treated (ATT) is:

$$ATT = E[Y(1) - Y(0) \mid T = 1] = E[Y(1) \mid T = 1] - E[Y(0) \mid T = 1],$$

where $E[Y(0) \mid T = 1]$ is inferred from matched controls. For our study, PSM balances segments on pre-construction traits, 'mimicking' randomization to isolate construction’s effect. This is vital since construction isn’t randomly assigned (i.e. busier roads might be more prone to work). We’ll use nearest-neighbor matching and assess balance with standardized differences. PSM excels at handling multiple covariates, fitting our US Road Construction and Closures dataset, but requires strong assumption (CIA) and sufficient overlap in $PSM(X)$ (common support).

## 2.3 Difference-in-Differences

Difference-in-Differences (DiD) is a quasi-experimental approach that estimates causal effects by comparing outcome changes over time between treated (roads with construction) and control (roads without) groups. It requires the parallel trends assumption: without treatment, accident rate differences between groups would stay constant over time. The model is:

$$Y_{it} = \beta_0 + \beta_1 T_i + \beta_2 Post_t + \beta_3 (T_i \times Post_t) + \epsilon_{it},$$

where $Y_{it}$ is the accident rate for segment $i$ at time $t$, $T_i$ is the treatment dummy (1 for construction, 0 otherwise), $Post_t$ is the time dummy (1 post-construction, 0 pre), and $\beta_3$ is the DiD estimator. This is calculated as:

$$\beta_3 = (E[Y_{T=1,Post}] - E[Y_{T=1,Pre}]) - (E[Y_{T=0,Post}] - E[Y_{T=0,Pre}]),$$

capturing the treatment effect by differencing pre-post changes across groups.

In our context, DiD compares accident rates before and after construction starts, between treated and control segments, controlling for time-invariant confounders and shared trends. We’ll verify parallel trends by plotting pre-treatment accident rates and testing statistically. Challenges include potential trend violations or spillovers (i.e. traffic rerouting to control roads), but DiD’s strength lies in exploiting time variation, making it ideal for tracing construction’s impact on accident trajectories.

# Part 3 - Implementation

## 3.0 Load Packages

Start by loading the packages required to run this section.

```{r Packages}
library(data.table)   # For faster data manipulation
library(tidyverse)    # For data manipulation and visualization
library(sf)           # For spatial data handling
library(leaflet)      # For interactive maps
library(leaflet.extras) # For additional leaflet features
library(mapview)      # For easier map visualization
library(tmap)         # For thematic maps
library(tigris)       # For US road networks
library(future)       # For parallel processing
library(future.apply) # For parallel processing with apply functions
library(sf)           # For spatial data handling
library(RPostgres)    # For PostgreSQL connection
library(DBI)          # For database interface

# Create directories if they don't exist
if (!dir.exists("./data/tigris")) {
  dir.create("./data/tigris", recursive = TRUE)
}

# Set custom cache directory (optional)
options(tigris_cache_dir = "./data/tigris")
# Configure tigris to use caching
options(tigris_use_cache = TRUE)
```

## 3.1 Descriptive Maps

### a. Static Maps

```{r Static Maps}
# Static map for Accidents with year coloring
# First, store your plot in a variable
ca_plot <- ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = df.acc.sf %>% filter(State == "CA"), 
          aes(color = factor(year)), alpha = 0.7, size = 0.5) +
  scale_color_viridis_d(name = "Year") +  # Use viridis color palette for discrete values
  theme_minimal() +
  labs(title = "US Road Accidents by Year (2016-2021)") +
  theme(legend.position = "bottom")
print(ca_plot)
# Then save it using ggsave
ggsave(filename = "imgs/california_accidents.png", 
       plot = ca_plot,
       width = 10, # width in inches
       height = 8, # height in inches
       dpi = 300)  # resolution

# Construction sites map with year coloring
# First, store our plot in a variable
ca_const_plot <- ggplot() +
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  geom_sf(data = df.const.sf %>% filter(State == "CA"), 
          aes(color = factor(year)), alpha = 0.7, size = 0.5) +
  scale_color_viridis_d(name = "Year") +  # Use viridis color palette for discrete values
  theme_minimal() +
  labs(title = "US Construction Sites by Year (2016-2021)") +
  theme(legend.position = "bottom")
print(ca_const_plot)
# Then save it using ggsave
ggsave(filename = "imgs/california_construction.png", 
       plot = ca_const_plot,
       width = 10, # width in inches
       height = 8, # height in inches
       dpi = 300)  # resolution
```


### b. Heatmaps

#### (i) Static

```{r Heatmap static}
# Create heatmap of accidents using ggplot2
# Modified accident heatmap with transparency for low density areas
accident_heatmap <- ggplot() +
  # Add California roads as base layer
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  # Create density heatmap with transparency threshold
  stat_density_2d(
    data = df.acc %>% filter(!is.na(Start_Lat) & !is.na(Start_Lng)),
    aes(x = Start_Lng, y = Start_Lat, fill = after_stat(density), alpha = after_stat(density)),
    geom = "tile", 
    contour = FALSE,
    h = 0.1,
    n = 200
  ) +
  # Set alpha scaling with a minimum threshold
  scale_alpha_continuous(range = c(0, 0.9), guide = "none") +
  # Use color scheme similar to leaflet's default heatmap
  scale_fill_gradientn(
    colors = c("#0000FF", "#00FFFF", "#00FF00", "#FFFF00", "#FF0000"),
    name = "Accident\nDensity"
  ) +
  coord_sf(
    xlim = c(-124.5, -114.5),
    ylim = c(32.5, 42.5)
  ) +
  theme_minimal() +
  labs(
    title = "Accident Density Heatmap in California (2016-2021)",
    x = NULL,
    y = NULL
  ) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold")
  )
# Display the plot
print(accident_heatmap)

ggsave(filename = "imgs/accident_heatmap.png", 
       plot = accident_heatmap,
       width = 10, # width in inches
       height = 8, # height in inches
       dpi = 300)  # resolution

```

```{r Construction Heatmap static}
# Create heatmap of construction sites using ggplot2
construction_heatmap <- ggplot() +
  # Add California roads as base layer
  geom_sf(data = ca_roads, color = "gray80", size = 0.1) +
  # Create density heatmap
  stat_density_2d(
    data = df.const %>% filter(!is.na(Start_Lat) & !is.na(Start_Lng)),
    aes(x = Start_Lng, y = Start_Lat, fill = after_stat(density)),
    geom = "tile", 
    contour = FALSE,
    alpha = 0.7,
    h = 0.1,  # Bandwidth
    n = 200   # Resolution
  ) +
  # Set alpha scaling with a minimum threshold
  scale_alpha_continuous(range = c(0, 0.9), guide = "none") +
  # Use color scheme matching your leaflet construction map
  scale_fill_gradientn(
    colors = c("yellow", "orange", "red"),
    name = "Construction\nDensity"
  ) +
  # Set appropriate boundaries for California
  coord_sf(
    xlim = c(-124.5, -114.5),
    ylim = c(32.5, 42.5)
  ) +
  theme_minimal() +
  labs(
    title = "Construction Density Heatmap in California (2016-2021)",
    x = NULL,
    y = NULL
  ) +
  theme(
    legend.position = "right",
    plot.title = element_text(face = "bold")
  )
# Display the plot
print(construction_heatmap)
```


#### (ii) Interactive


```{r Interactive Maps I}
# For interactive maps (often better for large datasets)
# Accidents map
accident_map <- leaflet() %>%
  addTiles() %>%
  addHeatmap(data = df.acc.sf, 
             intensity = ~1,
             radius = 8, 
             blur = 10) %>%
  setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>% # Center to CA
  setMaxBounds(lng1 = -124.6, lat1 = 42.0,    # Northwest corner of CA
               lng2 = -114.1, lat2 = 32.5)    # Southeast corner of CA

# Display maps
accident_map
```


```{r Interactive Maps II}
# Construction sites map with California focus and bounds constraints
construction_map <- leaflet() %>%
  addTiles() %>%
  addHeatmap(data = df.const.sf, 
             intensity = ~1,
             radius = 8, 
             blur = 10,
             gradient = c("yellow", "orange", "red")) %>%
  setView(lng = -119.4179, lat = 36.7783, zoom = 6) %>%  # Center on California
  setMaxBounds(lng1 = -124.6, lat1 = 42.0,    # Northwest corner of CA
               lng2 = -114.1, lat2 = 32.5)    # Southeast corner of CA

# Display the map
construction_map
```

<hr>



## 3.2 Propensity Score Matching

## 3.3 Difference-in-Differences

# Part 4 - Discussion

## 4.1 Results

## 4.2 Limitations

## 4.3 Improvements